---
title: "Introduction to irtlink"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to irtlink}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

irtlink estimates trends in the mean and standard deviation of
latent ability distributions across the time points of educational
assessments, and ability distribution differences across linked
groups, with unidimensional item response theory (IRT) models.
Under differential item functioning (DIF), the linked estimates
depend on the selected common items. The package estimates this
linking error together with the standard error.

```{r}
library(irtlink)
```

## Simulated data

`sim_trend_data()` generates dichotomous responses for several
groups with a known trend. `overlap` sets the proportion of common
items between successive groups.

```{r}
set.seed(123)
d <- sim_trend_data(n_groups = 3, N = 1000, I = 20, overlap = 1,
                    mu = c(0, 0.3, 0.5), sigma = c(1, 1.2, 1.5),
                    dif = "random", dif_pct = 1, dif_sd = 0.3)
str(d, max.level = 1)
```

## Calibration

`calibrate()` fits one IRT model per group on the internal marginal
maximum likelihood EM engine. The default is the 2PL with
`calibration = "separate"`. The alternatives are `"concurrent"`,
`"fixed"` (fixed item parameter calibration, FIPC), and
`"regularized"` (regularized calibration with a smooth Bayesian
information criterion, SBIC).

```{r}
cal <- calibrate(d)
cal
```

## Linking

`link_groups()` places the separately calibrated groups on one
common scale. `approach` selects the chain or the joint composition.
`method` selects mean-mean, mean-geometric mean, Haberman, Haebara,
or Stocking-Lord linking.

```{r}
link <- link_groups(cal, approach = "chain", method = "mgm")
link$trend
```

`ref` selects the reference group of the reported trend.

```{r}
link_groups(cal, approach = "chain", method = "mgm", ref = 2)$trend
```

## Linking error

`linking_error()` estimates the linking error of the linked means
and standard deviations. `method` selects the delete-one item
jackknife, the approximate jackknife, or a sandwich estimator.

```{r}
linking_error(link, method = "jackknife")$le
linking_error(link, method = "sandwich_osw")$le
```

## Full pipeline

`estimate_trend()` runs calibration, linking, and error estimation
in one call. The bias-corrected methods decompose the uncertainty
into standard error (`se`), linking error (`le`, `le_bc`), and total
error (`te`, `te_bc`).

```{r}
tr <- estimate_trend(d, calibration = "separate", approach = "chain",
                     link = "mgm", le = "sandwich_osw_bc")
round(tr$trend[, c("group", "mu", "sigma", "se", "le", "le_bc",
                   "te", "te_bc")], 3)
```

For dependent person samples, `dependent_vcov()` provides the joint
item parameter covariance for the `le_vcov` argument.

## DIF detection

`detect_dif()` flags items by root mean square deviation (RMSD)
statistics or by per-item likelihood-ratio tests, optionally with
iterative purification.

```{r}
dif <- detect_dif(cal, method = "rmsd")
dif$flagged
head(dif$rmsd)
```

The function groups and the references are listed in `?irtlink`.
